---
title: "index"
author: "Nischal Shakya"
date: "2/1/2020"
output: 
    html_document: 
      fig_caption: yes
      fig_width: 8
      highlight: haddock
      keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

remove(list = ls())

library('dplyr')

library('ggplot2')

library('rmarkdown')

source('./utils.R')

setwd('/home/nischal/nischalshakya15.github.io/work-related-learning')

javaProgrammingOneDf <- read.csv('data-sets/JavaProgrammingOne.csv')

javaTwoDf <- read.csv('data-sets/JavaProgrammingTwo.csv')

javaWebDf <- read.csv('data-sets/JavaWebProgramming.csv')
```

## Java Programming One 
This section shows the information about Java Programming One.
```{r Analysis of Java Programming One, echo=FALSE}
colnames(javaProgrammingOneDf)

javaProgrammingOneDf <- javaProgrammingOneDf %>% rename(
                        Tutorial.Full.Marks = Full.Marks, Tutorial.Marks.Obtained = Marks.Obtained,
                        Group.Project.Full.Marks = Full.Marks.1, Group.Project.Obtained.Marks = Marks.Obtained.1,
                        Assignment.Full.Marks = Full.Marks.2, Assignment.Marks.Obtained = Marks.Obtained.2,
                        Mid.Term.Full.Marks = Full.Marks.3, Mid.Term.Marks.Obtained = Marks.Obtained.3, Mid.Term.FifteenPercent.Marks = Marks.Obtained.4,
                        Lab.Test.Full.Marks = Full.Marks.4, Lab.Test.Marks.Obtained = Marks.Obtained.5, Lab.Test.FifteenPercent.Marks = Marks.Obtained.6,
                        FinalExam.Full.Marks = Full.Marks.5, Final.Exam.Marks.Obtained = Marks.Obtained.7, Final.Exam.FortyPercent.Marks = Marks.Obtained..40.)

colnames(javaProgrammingOneDf)
```

Since, we are grouping student on basiscs of grade we have to filter unique grade.
```{r Unique grades of Java Programming One, echo=FALSE}
uniqueGradeJavaProgrammingOne <- getUniqueAttribute(javaProgrammingOneDf$Grade)
print(uniqueGradeJavaProgrammingOne)
```

General inforamtion about variables
```{r General inforamtion about data of java programming one, echo=FALSE}
str(javaProgrammingOneDf)
```

Now, what we want to know is total number number of student on basics of grade.
```{r Counting totoal number of student of java programming one onn basics of grade, echo=FALSE}
javaProgrammingOneCountDf <- data.frame()

for (u in uniqueGradeJavaProgrammingOne) {
  javaProgrammingOneCountDf <- rbind(javaProgrammingOneCountDf, 
                      data.frame(
                        NoOfStudent = javaProgrammingOneDf %>% filter(Grade == u) %>% count(Grade)
                      ))
}

javaProgrammingOneCountDf <- sortInAscendingOrder(javaProgrammingOneCountDf, javaProgrammingOneCountDf$NoOfStudent.n)

paged_table(javaProgrammingOneCountDf)

```

Plot bargraph to know the total number of student on basics of grade.
```{r Java programming one bargraph, echo=FALSE}
plotBarGraph(df = javaProgrammingOneCountDf, x = javaProgrammingOneCountDf$NoOfStudent.Grade, 
             y = javaProgrammingOneCountDf$NoOfStudent.n, label = javaProgrammingOneCountDf$NoOfStudent.n, 
             title = 'No of student on basics of grade',
             xlab = 'Grade', ylab = 'Total No of Student')
```

Top 5 student of java programming one
```{r Java programming one top five student, echo = FALSE}
javaProgrammingOneTopFive <- findTopTen(javaProgrammingOneDf, javaProgrammingOneDf$Grade.Point, top = 5)
javaProgrammingOneTopFive <- sortInAscendingOrder(javaProgrammingOneTopFive, javaProgrammingOneTopFive$Grade.Point)
paged_table(javaProgrammingOneTopFive %>% select(Name, Total, Grade, Grade.Point))
```

Total number of student attending java programming one, summary and standard deviation
```{r Total number of student attending java programming one, echo = FALSE}
totalStudentInJavaOne <- javaProgrammingOneCountDf %>% select(NoOfStudent.n) %>% sum()
print(totalStudentInJavaOne)

summary(javaProgrammingOneDf$Total)
sd(javaProgrammingOneDf$Total)
```

## Java Programing Two 
This section shows the information about java programming two.
```{r Analysis of java programming one, echo = FALSE} 
colnames(javaTwoDf)

javaTwoDf <- javaTwoDf %>% rename(
             Mid.Term.Full.Marks = Full.Marks, Mid.Term.Marks.Obtained = Score, Mid.Term.Fifteen.Percent.Marks = Marks.Obtained..15..,
             Lab.Test.Full.Marks = Full.Marks.1, Lab.Test.Marks.Obtained = Marks.Obtained..15...1,
             Assignment.Full.Marks = Full.Marks.2, Assignment.Marks.Obtained = Marks.Obtained..25..,
             Tutorial.Full.Marks = Full.Marks.3, Tutorial.Marks.Obtained = Marks.Obtained..5..,
             Internal.Marks = Internal.Marks..60.., 
             Final.Exam.Full.Marks = Full.Marks.4, Final.Exam.Marks.Obtained = Score.1, Final.Exam.FortyPercent.Marks = Marks.Obtained..40..,
             Total = Total..100.., 
             Grade = Grade.after.40..rule
)

colnames(javaTwoDf)
```

Filter unique grades 
``` {r Java two unique grades}
uniqueGradeJavaTwo <- getUniqueAttribute(javaTwoDf$Grade)

print(uniqueGradeJavaTwo)

```
Total number of student on basics of grade. 
``` {r Java two total number of student on basics of grade, echo = FALSE}
javaTwoCountDf <- data.frame()

str(javaTwoDf)

for (u in uniqueGradeJavaTwo) {
  javaTwoCountDf <- rbind(javaTwoCountDf, 
                                     data.frame(
                                       NoOfStudent = javaTwoDf %>% filter(Grade == u) %>% count(Grade)
                                     ))
}
javaTwoCountDf <- sortInAscendingOrder(javaTwoCountDf, javaTwoCountDf$NoOfStudent.n)

paged_table(javaTwoCountDf)
```
Bar graph showing total number of student on basics of grade 
```{r Java two bargraph showing student on basics of grade, echo = FALSE} 
plotBarGraph(df = javaTwoCountDf, x = javaTwoCountDf$NoOfStudent.Grade, 
             y = javaTwoCountDf$NoOfStudent.n, label = javaTwoCountDf$NoOfStudent.n, 
             title = 'No of student on basics of grade',
             xlab = 'Grade', ylab = 'Total No of Student')
```
Top 5 student of java programming one 
```{r Java two top 5 student, echo = FALSE} 
javaTwoTopFive <- findTopTen(javaTwoDf, javaTwoDf$Grade.Point, top = 5)
javaTwoTopFive <- sortInAscendingOrder(javaTwoTopFive, javaTwoTopFive$Grade.Point)
paged_table(javaTwoTopFive %>% select(Name, Total, Grade, Grade.Point))

```

Total number of student, summary and standard deviation 
```{r Java two summary, echo = FALSE} 
totalStudentInJavaTwo <- javaTwoCountDf %>% select(NoOfStudent.n) %>% sum()
print(totalStudentInJavaTwo)
summary(javaTwoDf$Total)
sd(javaTwoDf$Total)
``` 
