---
title: "Student Data analysis of Java Programming"
author: Nischal Shakya
date: 8/23/20
output: html_document
---

<style type="text/css">
body {
    font-size: 13px;
}

h1 {
  font-size: 28px;
  color: DarkBlue;
}

h2 {
  font-size: 22px;
  color: DarkBlue;
}

th .pagedtable-header-type {
  display: none;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

remove(list = ls())

library('knitr')

library('dplyr')

library('ggplot2')

library('ggpubr')

library('fpc')

library('psych')

library('cluster')

library('rmarkdown')

source('utils.R')

setwd('/home/nischal/repository/personal/R/nischalshakya15.github.io/work-related-learning')

## Java One

# Start read
javaOneDf <- read.csv('data-sets/re-modified/JavaProgrammingOne.csv')
# End

javaOneDf <- mutateRank(javaOneDf, javaOneDf$Total)

# Start Get unique Grade and Rank
javaOneUniqueGrades <- getUnique(javaOneDf$Grade)
javaOneRanks <- getUnique(javaOneDf$Rank)
# End

## End

## Java Two
javaTwoDf <- read.csv('data-sets/re-modified/JavaProgrammingTwo.csv')

javaTwoDf <- mutateRank(javaTwoDf, javaTwoDf$Total)

# Start Get unique Grade and Rank
javaTwoGrades <- getUnique(javaTwoDf$Grade)
javaTwoRanks <- getUnique(javaTwoDf$Rank)
# End

# Start Replace empty value with M and Female with F
i <- 1
for (row in rownames(javaTwoDf)) {
  isFemale <- javaTwoDf[row, "Gender"]
  if (isFemale == "") {
    javaTwoDf$Gender[i] <- "M"
  } else {
    javaTwoDf$Gender[i] <- "F"
  }
  i <- i + 1
}
# End

## End

## Start Java Web Programming
javaWebDf <- read.csv('data-sets/re-modified/JavaWebProgramming.csv')

javaWebDf <- mutateRank(javaWebDf, javaWebDf$Total)

# Start Get unique Grade and Rank
javaWebGrades <- getUnique(javaWebDf$Grade)
javaWebRanks <- getUnique(javaWebDf$Rank)
# End

# Start Replace empty value with M and Female with F
i <- 1
for (row in rownames(javaWebDf)) {
  isFemale <- javaWebDf[row, "Gender"]
  if (isFemale == "") {
    javaWebDf$Gender[i] <- "M"
  } else {
    javaWebDf$Gender[i] <- "F"
  }
  i <- i + 1
}
## End
```

# Java Programming Student Result Analysis and Visualization
This section describes the data analysis of the student of same batch who have given the exam. Referring to the Malaysian Standard, Java divided into 3 modules i.e. Java One, Java Two and JavaWeb.


# Barchart showing total number of students on basics of Grade
## JavaProgramming One
```{r JavaOne showing total number of student on basics of Grade, fig.cap='Fig 1:- JavaProgramming One Student on basics of Grade', fig.align='center', echo = FALSE}
# Start plot bargraph showing total no of student on basics of Grade
javaOneDfCount <- data.frame()
for (g in javaOneUniqueGrades) {
  javaOneDfCount <- rbind(javaOneDfCount,
                          data.frame(
                            countGrade(javaOneDf, g)
                          ))
}
javaOneDfCount <- sortInAscendingOrder(javaOneDfCount, javaOneDfCount$n)
plotBarGraph(df = javaOneDfCount, x = 'Grade', y = 'n',
             label = javaOneDfCount$n, xlab = 'Grade', ylab = 'Number of Student')
# End
```

## JavaProgrammingTwo
```{r JavaTwo showing total number of student on basics of Grader, fig.cap='Fig 2:- Java Two Student on basics of Grade', fig.align='center', echo = FALSE}
# Start Plot bargraph showing total no of students on basics of Grade
javaTwoGradeWiseDf <- data.frame()
for (g in javaTwoGrades) {
  javaTwoGradeWiseDf <- rbind(javaTwoGradeWiseDf,
                              data.frame(
                                countGrade(javaTwoDf, g)
                              ))
}
javaTwoGradeWiseDf <- sortInAscendingOrder(javaTwoGradeWiseDf, javaTwoGradeWiseDf$n)
plotBarGraph(df = javaTwoGradeWiseDf, x = 'Grade', y = 'n',
             label = javaTwoGradeWiseDf$n, xlab = 'Grade', ylab = 'Number of Student')
# End
```

## JavaWebProgramming
```{r}

```
```{r JavaWeb showing total number of student on basics of Rank, fig.cap='Fig 3:- Java Web Student on basics of Grade', fig.align='center', echo = FALSE}
# Start Plot bargraph showing total no of students on basics of Grade
javaWebGradeWiseDf <- data.frame()
for (g in javaWebGrades) {
  javaWebGradeWiseDf <- rbind(javaWebGradeWiseDf,
                              data.frame(
                                countGrade(javaWebDf, g)
                              ))
}
javaWebGradeWiseDf <- sortInAscendingOrder(javaWebGradeWiseDf, javaWebGradeWiseDf$n)
plotBarGraph(df = javaWebGradeWiseDf, x = 'Grade', y = 'n',
             label = javaWebGradeWiseDf$n, xlab = 'Grade', ylab = 'Number of Student')
# End
```


# Barchart showing total number of student on basics of Rank
## JavaProgrammingOne
```{r JavaProgrammingOne showing total number of student on basics of Rank, fig.cap='Fig 4:- JavaProgramming One Student on basics of Rank', fig.align='center', echo = FALSE}
# Start plot bargraph showing total no of students on basics of Rank
javaOneRankWiseDf <- data.frame()
for (r in javaOneRanks) {
  javaOneRankWiseDf <- rbind(javaOneRankWiseDf,
                             data.frame(
                               getCountRank(javaOneDf, r)
                             ))
}
javaOneRankWiseDf <- sortInAscendingOrder(javaOneRankWiseDf, javaOneRankWiseDf$n)
plotBarGraph(df = javaOneRankWiseDf, x = 'Rank', y = 'n', label = javaOneRankWiseDf$n, xlab = 'Rank', ylab = 'Number of student')
# End
```

## JavaProgramingTwo
```{r JavaProgrammingTwo showing total number of student on basics of Rank, fig.cap='Fig 5:- JavaProgrammingTwo Student on basics of Rank', fig.align='center', echo = FALSE}
# Start Plot bargraph showing total no of students on basics of Rank
javaTwoRankWiseDf <- data.frame()
for (r in javaTwoRanks) {
  javaTwoRankWiseDf <- rbind(javaTwoRankWiseDf,
                             data.frame(
                               getCountRank(javaTwoDf, r)
                             ))
}
javaTwoRankWiseDf <- sortInAscendingOrder(javaTwoRankWiseDf, javaTwoRankWiseDf$n)
plotBarGraph(df = javaTwoRankWiseDf, x = 'Rank', y = 'n', label = javaTwoRankWiseDf$n, xlab = 'Rank', ylab = 'Number of student')
#End
```

## JavaWebProgramming
```{r JavaWebProgramming showing total number of student on basics of Rank, fig.cap='Fig 6:- JavaWebProgramming Student on basics of Rank', fig.align='center', echo = FALSE}
# Start Plot bargraph showing total no of students on basics of rank
javaWebRankWiseDf <- data.frame()
for (r in javaWebRanks) {
  javaWebRankWiseDf <- rbind(javaWebRankWiseDf,
                             data.frame(
                               getCountRank(javaWebDf, r)
                             ))
}
javaWebRankWiseDf <- sortInAscendingOrder(javaWebRankWiseDf, javaWebRankWiseDf$n)
plotBarGraph(df = javaWebRankWiseDf, x = 'Rank', y = 'n', label = javaWebRankWiseDf$n, xlab = 'Rank', ylab = 'Number of student')
# End
```


# Stacked barchart showing male and female on basics of Rank
## Java Programming One
```{r JavaProgrammingOne showing total number of male and female on basics of Rank, fig.cap='Fig 7:- JavaProgrammingOne Student on basics of Rank', fig.align='center', echo = FALSE}
# Start plot stackbar chart showing no of males and female on basics of Rank
javaOneRankWiseGenderDf <- data.frame()
javaOneRankWiseBarChartDf <- data.frame()
for (r in javaOneRanks) {
  rankFilter <- javaOneDf %>% filter(Rank == r)
  femaleRankFilter <- rankFilter %>%
    filter(Gender == "F") %>%
    count(Gender)
  if (length(femaleRankFilter$n) == 0) {
    femaleRankFilter <- data.frame(Gender = "F", n = 0)
  }
  javaOneRankWiseGenderDf <- rbind(javaOneRankWiseGenderDf,
                                   data.frame(
                                     Rank = r,
                                     Male = rankFilter %>%
                                       filter(Gender == "M") %>%
                                       count(Gender),
                                     Female = femaleRankFilter
                                   ))
}

javaOneRankWiseGenderDf <- javaOneRankWiseGenderDf %>% select(-c(Male.Gender, Female.Gender))

for (r in javaOneRanks) {
  filter <- javaOneRankWiseGenderDf %>% filter(Rank == r)
  javaOneRankWiseBarChartDf <- rbind(javaOneRankWiseBarChartDf,
                                     data.frame(
                                       Rank = rep(r, 2),
                                       Gender = c(rep('Male', 1), rep('Female', 1)),
                                       NoOfStudent = c(filter$Male.n, filter$Female.n)
                                     ))
}

plotStackBar(javaOneRankWiseBarChartDf, x = 'Rank', y = 'NoOfStudent',
             fill = 'Gender', color = 'Gender', palette = 'uchiago',
             label = javaOneRankWiseBarChartDf$NoOfStudent, xlab = 'Rank', ylab = 'No of student')
# End
```

## JavaProgrammingTwo
```{r JavaProgrammingTwo showing total number of male and female on basics of Rank, fig.cap='Fig 8:- JavaProgrammingOne Student on basics of Rank', fig.align='center', echo = FALSE}
# Start plot stackbar chart showing no of males and female on basics of Rank
javaOneRankWiseGenderDf <- data.frame()
javaOneRankWiseBarChartDf <- data.frame()
for (r in javaOneRanks) {
  rankFilter <- javaOneDf %>% filter(Rank == r)
  femaleRankFilter <- rankFilter %>%
    filter(Gender == "F") %>%
    count(Gender)
  if (length(femaleRankFilter$n) == 0) {
    femaleRankFilter <- data.frame(Gender = "F", n = 0)
  }
  javaOneRankWiseGenderDf <- rbind(javaOneRankWiseGenderDf,
                                   data.frame(
                                     Rank = r,
                                     Male = rankFilter %>%
                                       filter(Gender == "M") %>%
                                       count(Gender),
                                     Female = femaleRankFilter
                                   ))
}

javaOneRankWiseGenderDf <- javaOneRankWiseGenderDf %>% select(-c(Male.Gender, Female.Gender))

for (r in javaOneRanks) {
  filter <- javaOneRankWiseGenderDf %>% filter(Rank == r)
  javaOneRankWiseBarChartDf <- rbind(javaOneRankWiseBarChartDf,
                                     data.frame(
                                       Rank = rep(r, 2),
                                       Gender = c(rep('Male', 1), rep('Female', 1)),
                                       NoOfStudent = c(filter$Male.n, filter$Female.n)
                                     ))
}

plotStackBar(javaOneRankWiseBarChartDf, x = 'Rank', y = 'NoOfStudent',
             fill = 'Gender', color = 'Gender', palette = 'uchiago',
             label = javaOneRankWiseBarChartDf$NoOfStudent, xlab = 'Rank', ylab = 'No of student')
# End
```

## JavaWebProgramming
```{r JavaWebProgramming showing total number of male and female on basics of Rank, fig.cap='Fig 9:- JavaProgrammingOne Student on basics of Rank', fig.align='center', echo = FALSE}
# Start stack bar chart showing no of male and female on basics of Rank
javaWebRankWiseGenderDf <- data.frame()

javaWebRankWiseBarChartDf <- data.frame()

for (r in javaWebRanks) {
  rankFilter <- javaWebDf %>% filter(Rank == r)
  maleRankFilter <- rankFilter %>%
    filter(Gender == "M") %>%
    count(Gender)
  femaleRankFilter <- rankFilter %>%
    filter(Gender == "F") %>%
    count(Gender)
  if (length(femaleRankFilter$n) == 0) {
    femaleRankFilter <- data.frame(Gender = "F", n = 0)
  }
  if (length(maleRankFilter$n) == 0) {
    maleRankFilter <- data.frame(Gender = "M", n = 0)
  }
  javaWebRankWiseGenderDf <- rbind(javaWebRankWiseGenderDf,
                                   data.frame(
                                     Rank = r,
                                     Male = maleRankFilter,
                                     Female = femaleRankFilter
                                   ))
}

javaWebRankWiseGenderDf <- javaWebRankWiseGenderDf %>% select(-c(Male.Gender, Female.Gender))

for (r in javaWebRanks) {
  filter <- javaWebRankWiseGenderDf %>% filter(Rank == r)
  javaWebRankWiseBarChartDf <- rbind(javaWebRankWiseBarChartDf,
                                     data.frame(
                                       Rank = rep(r, 2),
                                       Gender = c(rep('Male', 1), rep('Female', 1)),
                                       NoOfStudent = c(filter$Male.n, filter$Female.n)
                                     ))
}

plotStackBar(javaWebRankWiseBarChartDf, x = 'Rank', y = 'NoOfStudent',
             fill = 'Gender', color = 'Gender', palette = 'uchiago',
             label = javaWebRankWiseBarChartDf$NoOfStudent, xlab = 'Rank', ylab = 'No of student')

# End
```

# Top 2 students from each Rank including one Male and Female.
## JavaProgrammingOne
```{r JavaOne top two student from each Rank, echo = FALSE}
javaOneTwoExcellentStudent <- javaOneDf %>%
  filter(Rank == 'Excellent') %>%
  top_n(2, Total)

javaOneTwoVeryGoodStudent <- bind_rows(javaOneDf %>%
                                         filter(Rank == 'Very Good' & Gender == 'M') %>%
                                         top_n(1, Total),
                                       javaOneDf %>%
                                         filter(Rank == 'Very Good' & Gender == 'F') %>%
                                         top_n(1, Total)
)

javaOneTwoGoodStudent <- bind_rows(javaOneDf %>%
                                     filter(Rank == 'Good' & Gender == 'M') %>%
                                     top_n(1, Total),
                                   javaOneDf %>%
                                     filter(Rank == 'Good' & Gender == 'F') %>%
                                     top_n(1, Total))

javaOneTwoStatisfactoryStudent <- bind_rows(javaOneDf %>%
                                              filter(Rank == 'Statisfactory' & Gender == 'M') %>%
                                              top_n(1, Total),
                                            javaOneDf %>%
                                              filter(Rank == 'Statisfactory' & Gender == 'F') %>%
                                              top_n(1, Total))
# End

# Start Gathering General Information
javaOneTopTwoStudents <- bind_rows(javaOneTwoExcellentStudent, javaOneTwoVeryGoodStudent, javaOneTwoGoodStudent, javaOneTwoStatisfactoryStudent)

javaOneTopTwoStudents <- javaOneTopTwoStudents %>% select(Name, Gender, Assignment.Marks.Obtained, Mid.Term.Fifteen.Percent.Marks.Obtained, Lab.Test.Fifteen.Percent.Marks.Obtained, Final.Exam.Forty.Percent.Marks, Total, Rank)
# End

paged_table(javaOneTopTwoStudents)
```

```{r JavaTwo top two student from each Rank, echo = FALSE}
# Start Pick up two student from each Rank
javaTwoTwoExcellentStudent <- javaTwoDf %>%
  filter(Rank == 'Excellent') %>%
  top_n(2, Total)

javaTwoTwoVeryGoodStudent <- bind_rows(javaTwoDf %>%
                                         filter(Rank == 'Very Good' & Gender == 'M') %>%
                                         top_n(1, Total),
                                       javaTwoDf %>%
                                         filter(Rank == 'Very Good' & Gender == 'F') %>%
                                         top_n(1, Total)
)

javaTwoTwoGoodStudent <- bind_rows(javaTwoDf %>%
                                     filter(Rank == 'Good' & Gender == 'M') %>%
                                     top_n(1, Total),
                                   javaTwoDf %>%
                                     filter(Rank == 'Good' & Gender == 'F') %>%
                                     top_n(1, Total))

javaTwoTwoStatisfactoryStudent <- bind_rows(javaTwoDf %>%
                                              filter(Rank == 'Statisfactory' & Gender == 'M') %>%
                                              top_n(1, Total),
                                            javaTwoDf %>%
                                              filter(Rank == 'Statisfactory' & Gender == 'F') %>%
                                              top_n(1, Total))
# End

# Start Gathering General Information
javaTwoStudents <- bind_rows(javaTwoTwoExcellentStudent, javaTwoTwoVeryGoodStudent, javaTwoTwoGoodStudent, javaTwoTwoStatisfactoryStudent)
javaTwoStudents <- javaTwoStudents %>% select(Name, Gender, Assignment.Marks.Obtained, Mid.Term.Fifteen.Percent.Marks, Lab.Test.Marks.Obtained, Final.Exam.FortyPercent.Marks, Total, Rank)
# End
paged_table(javaTwoStudents)
```

```{r JavaWeb Top two student, echo = FALSE}
# Start Pick up two student from each Rank
javaWebTwoExcellentStudent <- javaWebDf %>%
  filter(Rank == 'Excellent') %>%
  top_n(2, Total)

javaWebTwoVeryGoodStudent <- bind_rows(javaWebDf %>%
                                         filter(Rank == 'Very Good' & Gender == 'M') %>%
                                         top_n(1, Total),
                                       javaWebDf %>%
                                         filter(Rank == 'Very Good' & Gender == 'F') %>%
                                         top_n(1, Total)
)

javaWebTwoGoodStudent <- bind_rows(javaWebDf %>%
                                     filter(Rank == 'Good' & Gender == 'M') %>%
                                     top_n(1, Total),
                                   javaWebDf %>%
                                     filter(Rank == 'Good' & Gender == 'F') %>%
                                     top_n(1, Total))

javaWebTwoStatisfactoryStudent <- bind_rows(javaWebDf %>%
                                              filter(Rank == 'Statisfactory' & Gender == 'M') %>%
                                              top_n(1, Total),
                                            javaWebDf %>%
                                              filter(Rank == 'Statisfactory' & Gender == 'F') %>%
                                              top_n(1, Total))
# End

# Start Gathering General Information
javaWebStudents <- bind_rows(javaWebTwoExcellentStudent, javaWebTwoVeryGoodStudent, javaWebTwoGoodStudent, javaWebTwoStatisfactoryStudent)
javaWebStudents <- javaWebStudents %>% select(Name, Gender, Assignment.Marks.Obtained, Mid.Term.Twenty.Percent.Marks, Lab.Test.Marks.Obtained, Final.Exam.Forty.Percent.Marks, Total, Rank)
# End

paged_table(javaWebStudents)
```
# Cluster analysis
## JavaProgrammingOne
```{r JavaOne K means clustering, fig.cap='Fig 10:- JavaOne cluster analysis', fig.align='center', echo = FALSE}
javaOneCluster <- javaOneDf %>% select(Name, Total, Rank)

javaOneCluster <- convertRankIntoNumber(javaOneCluster, javaOneCluster$Total)

javaOneClusterNum <- javaOneCluster[c('Total', 'Rank')]

javaOnePamk <- pamk(javaOneClusterNum, krange = 2:5, metric = 'manhattan')

javaOnePam <- pam(x = javaOneClusterNum, k = 6, metric = 'manhattan')

javaOnePamCluster <- rep('NA', length(javaOneCluster$Total))

javaOnePamCluster[javaOnePam$clustering == 1] <- 'Cluster 1'
javaOnePamCluster[javaOnePam$clustering == 2] <- 'Cluster 2'
javaOnePamCluster[javaOnePam$clustering == 3] <- 'Cluster 3'
javaOnePamCluster[javaOnePam$clustering == 4] <- 'Cluster 4'
javaOnePamCluster[javaOnePam$clustering == 5] <- 'Cluster 5'
javaOnePamCluster[javaOnePam$clustering == 6] <- 'Cluster 6'

javaOneCluster$Cluster <- javaOnePamCluster

javaOneCluster$Cluster <- factor(javaOneCluster$Cluster, levels = c('Cluster 1', 'Cluster 2', 'Cluster 3', 'Cluster 4', 'Cluster 5', 'Cluster 6'))

ggplot(javaOneCluster,
       aes(x = Rank,
           y = Total,
           color = Cluster)) +
  geom_point(size = 3) +
  xlab('Rank') +
  ylab('Total') +
  geom_jitter(width = 0.4, height = 0.4) +
  theme_bw()
#End
```

## JavaProgrammingTwo
```{r JavaTwo K means clustering, fig.cap='Fig 11:- JavaTwo cluster analysis', fig.align='center', echo = FALSE}
javaTwoCluster <- javaTwoDf %>% select(Name, Total, Rank)

javaTwoCluster <- convertRankIntoNumber(javaTwoCluster, javaTwoCluster$Total)

javaOneClusterNum <- javaTwoCluster[c('Total', 'Rank')]

javaOnePamk <- pamk(javaOneClusterNum, krange = 2:5, metric = 'manhattan')

javaTwoPam <- pam(x = javaOneClusterNum, k = 6, metric = 'manhattan')

javaTwoPamCluster <- rep('NA', length(javaTwoCluster$Total))

javaTwoPamCluster[javaTwoPam$clustering == 1] <- 'Cluster 1'
javaTwoPamCluster[javaTwoPam$clustering == 2] <- 'Cluster 2'
javaTwoPamCluster[javaTwoPam$clustering == 3] <- 'Cluster 3'
javaTwoPamCluster[javaTwoPam$clustering == 4] <- 'Cluster 4'
javaTwoPamCluster[javaTwoPam$clustering == 5] <- 'Cluster 5'
javaTwoPamCluster[javaTwoPam$clustering == 6] <- 'Cluster 6'

javaTwoCluster$Cluster <- javaTwoPamCluster

javaTwoCluster$Cluster <- factor(javaTwoCluster$Cluster, levels = c('Cluster 1', 'Cluster 2', 'Cluster 3', 'Cluster 4', 'Cluster 5', 'Cluster 6'))

ggplot(javaTwoCluster,
       aes(x = Rank,
           y = Total,
           color = Cluster)) +
  geom_point(size = 3) +
  xlab('Rank') +
  ylab('Total') +
  geom_jitter(width = 0.4, height = 0.4) +
  theme_bw()
```

## JavaWebProgramming
```{r JavaTwo cluster analysis,fig.cap='Fig 12:- JavaWeb cluster analysis', fig.align='center', echo=FALSE}
# K means Clustering Algorithm Start
javaWebCluster <- javaOneDf %>% select(Name, Total, Rank)

javaWebCluster <- convertRankIntoNumber(javaWebCluster, javaWebCluster$Total)

javaWebClusterNum <- javaWebCluster[c('Total', 'Rank')]

javaWebPamk <- pamk(javaWebClusterNum, krange = 2:5, metric = 'manhattan')

javaWebPam <- pam(x = javaWebClusterNum, k = 6, metric = 'manhattan')

javaWebPamCluster <- rep('NA', length(javaWebCluster$Total))

javaWebPamCluster[javaWebPam$clustering == 1] <- 'Cluster 1'
javaWebPamCluster[javaWebPam$clustering == 2] <- 'Cluster 2'
javaWebPamCluster[javaWebPam$clustering == 3] <- 'Cluster 3'
javaWebPamCluster[javaWebPam$clustering == 4] <- 'Cluster 4'
javaWebPamCluster[javaWebPam$clustering == 5] <- 'Cluster 5'
javaWebPamCluster[javaWebPam$clustering == 6] <- 'Cluster 6'

javaWebCluster$Cluster <- javaWebPamCluster

javaWebCluster$Cluster <- factor(javaWebCluster$Cluster, levels = c('Cluster 1', 'Cluster 2', 'Cluster 3', 'Cluster 4', 'Cluster 5', 'Cluster 6'))

ggplot(javaWebCluster,
       aes(x = Rank,
           y = Total,
           color = Cluster)) +
  geom_point(size = 3) +
  xlab('Rank') +
  ylab('Total') +
  geom_jitter(width = 0.4, height = 0.4) +
  theme_bw()
# End
```