---
title: "Data analysis and visualization"
author: "NischalShakya"
date: "December 14, 2019"
output: 
    html_document: 
      fig_caption: yes
      fig_width: 8
      highlight: haddock
      keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clearing enviroment variables
remove(list=ls())
# Using library dplyr for data manupulation
library('dplyr')
# Using library ggplot2 
library('ggplot2')
# Using library rmarkdown
library('rmarkdown')

# Set current working directory 
setwd('/home/nischal/nischalshakya15.github.io/r-assignment')

# Load data from csv file into data frame
data <- read.csv('input.csv', sep = ',', header = TRUE, stringsAsFactors = FALSE)

# Remove Zone Bagmati and Development Region beacuse it is redudant
data <- select(data, -c(Zone, Development.Region, School.Number, School.Code, ENGLISH, NEPALI, MATHS, SCIENCE, SOCIAL.STUDIES, HEALTH.POPULATION.AND.ENVIRONMENT))

# Calculate fail percentage
data <- data %>% mutate('FAIL.PERCENTAGE' = 100 - data$PASS.PERCENT)

districts <- c(unique(data$District))

totalNoOfStudentDataFrame <- data.frame()

hundredPercentPassPercentDataFrame <- data.frame()

failPercentDataFrame <- data.frame()
zeroDistinctionFailPercentDataFrame <- data.frame()
distinctionFailPercentDataFrame <- data.frame()
totalGradeOfZeroDistinctionFailPercentDataFrame <- data.frame()
totalGradeOfDistinctionFailPercentDataFrame <- data.frame()
totalGradeOfZeroDistinctionBarChartDf <- data.frame()
totalGradeOfDistinctionFailPercentBarChartDf <- data.frame()

schoolGradeOfHundredPassPercentWithDistinctionDataFrame <- data.frame()
schoolGradeOfHundredPassPercentWithoutDistinctionDataFrame <- data.frame()

for (d in districts) {
  
  filterDistrict <- filter(data, District == d)
  

  totalNoOfStudentDataFrame <- rbind(totalNoOfStudentDataFrame, data.frame(District = d,
      
                                                                           Distinction = filterDistrict %>% select(DISTINCTION) %>% sum(),
                                                                           FirstDivision = filterDistrict %>% select(FIRST.DIVISION) %>% sum(),
                                                                           SecondDivision = filterDistrict %>% select(SECOND.DIVISION) %>% sum(),
                                                                           ThirdDivision = filterDistrict %>% select(THIRD.DIVISION) %>% sum(),
                                                                           Pass = filterDistrict %>% select(PASS) %>% sum(),
                                                                           Fail = filterDistrict %>% select(FAIL) %>% sum(),
                                                                           Total = filterDistrict %>% select(PASS,FAIL) %>% sum()))
  
  hundredPercentPassPercentDataFrame <- rbind(hundredPercentPassPercentDataFrame, 
                                              filter(data, District == d & FAIL == 0 & PASS.PERCENT == max(PASS.PERCENT)))
  
  failPercentDataFrame <- rbind(failPercentDataFrame, filter(data, District == d & FAIL != 0 & FAIL.PERCENTAGE != 0.0))
  
  
  zeroDistinctionFailPercentDataFrame <- rbind(zeroDistinctionFailPercentDataFrame, filter(failPercentDataFrame, District == d & DISTINCTION == 0))
  
  filterDistrictOfZeroDistinctionFailPercentDataFrame <- filter(zeroDistinctionFailPercentDataFrame, District == d)
  
  totalGradeOfZeroDistinctionFailPercentDataFrame <- rbind(totalGradeOfZeroDistinctionFailPercentDataFrame,
                                                           data.frame(
                                                             District = d, 
                                                             Distinction = filterDistrictOfZeroDistinctionFailPercentDataFrame %>% select(DISTINCTION) %>% sum(),
                                                             FirstDivision = filterDistrictOfZeroDistinctionFailPercentDataFrame %>% select(FIRST.DIVISION) %>% sum(),
                                                             SecondDivision = filterDistrictOfZeroDistinctionFailPercentDataFrame %>% select(SECOND.DIVISION) %>% sum(),
                                                             ThirdDivision = filterDistrictOfZeroDistinctionFailPercentDataFrame %>% select(THIRD.DIVISION) %>% sum(),
                                                             Total = filterDistrictOfZeroDistinctionFailPercentDataFrame %>% select(DISTINCTION,FIRST.DIVISION,SECOND.DIVISION,THIRD.DIVISION) %>% sum()))
  
  totalGradeOfZeroDistinctionBarChartDf <- rbind(totalGradeOfZeroDistinctionBarChartDf,
                                                                   data.frame(
                                                                     District = c(rep(d, 3)),
                                                                     Grade = c(rep("FirstDivision", 1), rep("SecondDivision", 1), rep("ThirdDivision",1)),
                                                                     NoOfStudent = c(filterDistrictOfZeroDistinctionFailPercentDataFrame %>% select(FIRST.DIVISION) %>% sum(), 
                                                                                     filterDistrictOfZeroDistinctionFailPercentDataFrame %>% select(SECOND.DIVISION) %>% sum(),
                                                                                     filterDistrictOfZeroDistinctionFailPercentDataFrame %>% select(THIRD.DIVISION) %>% sum()
                                                                     )
                                                                   ))
  
  distinctionFailPercentDataFrame <- rbind(distinctionFailPercentDataFrame, filter(failPercentDataFrame, District == d & DISTINCTION != 0))
  
  filterDistrictOfDistinctionFailPercentDataFrame <- filter(distinctionFailPercentDataFrame, District == d)
  
  totalGradeOfDistinctionFailPercentDataFrame <- rbind(totalGradeOfDistinctionFailPercentDataFrame, 
                                                       data.frame(
                                                         District = d, 
                                                         Distinction = filterDistrictOfDistinctionFailPercentDataFrame %>% select(DISTINCTION) %>% sum(),
                                                         FirstDivision = filterDistrictOfDistinctionFailPercentDataFrame %>% select(FIRST.DIVISION) %>% sum(),
                                                         SecondDivision = filterDistrictOfDistinctionFailPercentDataFrame %>% select(SECOND.DIVISION) %>% sum(),
                                                         ThirdDivision = filterDistrictOfDistinctionFailPercentDataFrame %>% select(THIRD.DIVISION) %>% sum(),
                                                         Total = filterDistrictOfDistinctionFailPercentDataFrame %>% select(DISTINCTION,FIRST.DIVISION,SECOND.DIVISION,THIRD.DIVISION) %>% sum()))
  
  totalGradeOfDistinctionFailPercentBarChartDf  <- rbind(totalGradeOfDistinctionFailPercentBarChartDf,
                                                         data.frame(
                                                           District = c(rep(d, 4)),
                                                           Grade = c(rep("Distinction", 1), rep("FirstDivision", 1), rep("SecondDivison",1), rep("ThirdDivision",1)),
                                                           NoOfStudent = c(filterDistrictOfDistinctionFailPercentDataFrame %>% select(DISTINCTION) %>% sum(), 
                                                                           filterDistrictOfDistinctionFailPercentDataFrame %>% select(FIRST.DIVISION) %>% sum(),
                                                                           filterDistrictOfDistinctionFailPercentDataFrame %>% select(SECOND.DIVISION) %>% sum(),
                                                                           filterDistrictOfDistinctionFailPercentDataFrame %>% select(THIRD.DIVISION) %>% sum())
                                                         ))
    
  schoolGradeOfHundredPassPercentWithoutDistinctionDataFrame <- rbind(schoolGradeOfHundredPassPercentWithoutDistinctionDataFrame, 
                                                                   filter(hundredPercentPassPercentDataFrame, District == d & DISTINCTION == 0))
  
  schoolGradeOfHundredPassPercentWithDistinctionDataFrame <- rbind(schoolGradeOfHundredPassPercentWithDistinctionDataFrame,
                                                                      filter(hundredPercentPassPercentDataFrame, District == d & DISTINCTION != 0))
}
```

## School performance for the year 2063 BS

This dataset includes the school performance data for the year 2063 BS based on the average scores and pass percentage in School Leaving Certificate (SLC) examination of Bagmati zone, an examination marking the completion of grade 10.

## General summary of column used in data set
```{r general summary of data}
str(data)
```

## Filtering unique districts of Bagmati Zone

```{r filtering unique district of Bagmati Zone, echo = FALSE}
# Get unique districts from data frame
districts <- c(unique(data$District))
```
``` {r unique district, echo = FALSE}
print(districts)
```

## Bargraph showing total no of student on basics of district
```{r Bar graph showing total no of student on basics of district, fig.cap='Fig 1:- Total No of Student in each district', fig.align='c', echo = FALSE}
paged_table(totalNoOfStudentDataFrame)

# Plot bar graph showing total no of student in each district
ggplot(totalNoOfStudentDataFrame, aes(x = District, y = Total, label = Total)) + geom_bar(stat = 'identity', fill = 'steelblue', color = 'black') + geom_text(size = 3, vjust = -1) + labs(title = 'Total no of student in each district') + xlab('Districts') + ylab('Total No of Student')
```

## No of school basics of district and geographical region {.tabset .tabset-fade}

### District

``` {r No of school on basics of district, echo = FALSE}
countNoofSchoolOnBasiscOfDistrict <- count(data, data$District)
paged_table(countNoofSchoolOnBasiscOfDistrict)
```
```{r Donut showing total no of school on basics of district, echo=FALSE}
# load countNoOfSchoolOnBasicsOfDistrict in donut dataframe
donut <- data.frame(
  District=countNoofSchoolOnBasiscOfDistrict$`data$District`,
  Count=countNoofSchoolOnBasiscOfDistrict$n
)

# Compute percentages
donut$fraction <- donut$Count / sum(donut$Count)

# Compute the cumulative percentages (top of each rectangle)
donut$ymax <- cumsum(donut$fraction)

# Compute the bottom of each rectangle
donut$ymin <- c(0, head(donut$ymax, n=-1))

# Compute label position
donut$labelPosition <- (donut$ymax + donut$ymin) / 2

# Compute a good label
donut$label <- paste0(donut$Count)

# Create donut 
ggplot(donut, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=District)) +
  geom_rect() +
  geom_text( x=3.5, aes(y=labelPosition, label=label), size=3) + # x here controls label position (inner / outer)
  scale_fill_brewer(palette=3) +
  scale_color_brewer(palette=3) +
  coord_polar(theta="y") +
  xlim(c(1, 4)) +
  theme_void() +
  theme(legend.position = "right")
```

### Geographical region

``` {r No of school on basics of geographical region, echo = FALSE}
countNoofSchoolOnBasiscOfGeographicalRegion <- count(data, data$Geographical.Region)
paged_table(countNoofSchoolOnBasiscOfGeographicalRegion)
```
``` {r Bar graph showing total no student on baiscs of geopgraphica region, echo = FALSE}
ggplot(countNoofSchoolOnBasiscOfGeographicalRegion, aes(x = countNoofSchoolOnBasiscOfGeographicalRegion$`data$Geographical.Region`, y = countNoofSchoolOnBasiscOfGeographicalRegion$n, label = countNoofSchoolOnBasiscOfGeographicalRegion$n)) + geom_bar(stat = 'identity', fill = 'steelblue', color = 'black', width = 0.3) + geom_text(size = 3, vjust = -1) + labs(title = 'Total no of school on basics of geographical region') + xlab('Geographical region') + ylab('Total No of School')
````

## No of school having 100 percent pass result
```{r No of school having 100 percent pass result, echo = FALSE} 
# Count no of school having 100% pass result
paged_table(select(hundredPercentPassPercentDataFrame, District, Name.of.School), options = list(rows.print = 15))
noOfSchoolWithHundredPercentPass <- count(hundredPercentPassPercentDataFrame, hundredPercentPassPercentDataFrame$District)
# Plot bar graph showing total no of school with 100% pass result 
ggplot(noOfSchoolWithHundredPercentPass, aes(x = noOfSchoolWithHundredPercentPass$`hundredPercentPassPercentDataFrame$District`, y = noOfSchoolWithHundredPercentPass$n, label = noOfSchoolWithHundredPercentPass$n)) + geom_bar(stat = 'identity', fill = 'steelblue', color = 'black', width = 0.5) + geom_text(size = 3, vjust = -1) + labs(title = 'Total no of school with 100% pass result') + xlab('District') + ylab('Total No of School')
```

Here, we can see Rasuwa was only a district without 100 percent pass result. Hence, we can assume that there is lack of quality education in Rasuwa district.

## General summary of 100% pass result {.tabset .tabset-fade}
Despite of 100% pass result, it seems in some school there is zero distiction. 
So, this seciton will summarize the total no of student on basics of grade having zero distinction within 100% pass result and vice versa.

### Having Distinction 
```{r No of student on baiscs of grade within 100 percent pass result school scoring non-zero distinction, echo=FALSE} 
totalGradeOfHundredPassPercentWithDistinction <- data.frame()
totalGradeOfHundredPassPercentWithDistinctionBarchartDf <- data.frame()
districtsHundredPassPercentWithDistinction <- c(unique(schoolGradeOfHundredPassPercentWithDistinctionDataFrame$District))
for (d in districtsHundredPassPercentWithDistinction) {
  
  filterDistrict <- filter(schoolGradeOfHundredPassPercentWithDistinctionDataFrame, District == d)
  
  totalGradeOfHundredPassPercentWithDistinction <- rbind(totalGradeOfHundredPassPercentWithDistinction, data.frame(
                                                                           District = d, 
                                                                           Distinction = filterDistrict %>% select(DISTINCTION) %>% sum(),
                                                                           FirstDivision = filterDistrict %>% select(FIRST.DIVISION) %>% sum(),
                                                                           SecondDivision = filterDistrict %>% select(SECOND.DIVISION) %>% sum(),
                                                                           ThirdDivision = filterDistrict %>% select(THIRD.DIVISION) %>% sum(),
                                                                           Total = filterDistrict %>% select(DISTINCTION,FIRST.DIVISION,SECOND.DIVISION,THIRD.DIVISION) %>% sum()))

  totalGradeOfHundredPassPercentWithDistinctionBarchartDf <- rbind(totalGradeOfHundredPassPercentWithDistinctionBarchartDf,
                                                                   data.frame(
                                                                     District =  c(rep(d, 3)),
                                                                     Grade = c(rep("Distinction", 1), rep("FirstDivision", 1), rep("SecondDivison",1)),
                                                                     NoOfStudent = c(filterDistrict %>% select(DISTINCTION) %>% sum(), 
                                                                                     filterDistrict %>% select(FIRST.DIVISION) %>% sum(),
                                                                                     filterDistrict %>% select(SECOND.DIVISION) %>% sum()
                                                                     )
                                                                   ))
  
}

paged_table(totalGradeOfHundredPassPercentWithDistinction)
``` 
From above table, we can conclude that school having 100% pass rate with distinction the no of student acheiving the third division is Zero. Hence, we can remove the column third division while visualizing data. 
```{r Stack bar chart representing no of student on basics of grade, echo=FALSE, fig.cap='Fig 4:- Total No of Student in each district on basics of grade', fig.align='c'} 

ggplot(totalGradeOfHundredPassPercentWithDistinctionBarchartDf, 
       aes(fill=Grade, 
           y=NoOfStudent, 
           x=District,
           label=NoOfStudent)) + 
  geom_bar(position='dodge', stat='identity') +
  xlab('District') + 
  ylab('Total No of Student')
```
From above barchart, we can conclude that Kathmandu district had the maximum pass rate in comparision to another district. So we can assume that another district lacks the proper educational facilities as well as school.

### Without Having Distinction 
```{r No of student on baiscs of grade within 100 percent pass result school scoring with zero distinction, echo = FALSE} 
# Initialize an empty dataframe 
totalGradeOfHundredPassPercentWithoutDistinction <- data.frame()
totalGradeOfHundredPassPercentWithoutDistinctionBarchartDf <- data.frame()
# Filtering unique district of hundred pass percent without distinction 
districtsHundredPassPercentWithoutDistinction <- c(unique(schoolGradeOfHundredPassPercentWithoutDistinctionDataFrame$District))

for (d in districtsHundredPassPercentWithoutDistinction) {
  
  filterDistrict <- filter(schoolGradeOfHundredPassPercentWithoutDistinctionDataFrame, District == d)
  
  totalGradeOfHundredPassPercentWithoutDistinction <- rbind(totalGradeOfHundredPassPercentWithoutDistinction, data.frame(
                                                                           District = d, 
                                                                           Distinction = filterDistrict %>% select(DISTINCTION) %>% sum(),
                                                                           FirstDivision = filterDistrict %>% select(FIRST.DIVISION) %>% sum(),
                                                                           SecondDivision = filterDistrict %>% select(SECOND.DIVISION) %>% sum(),
                                                                           ThirdDivision = filterDistrict %>% select(THIRD.DIVISION) %>% sum(),
                                                                           Total = filterDistrict %>% select(DISTINCTION,FIRST.DIVISION,SECOND.DIVISION,THIRD.DIVISION) %>% sum()))
  
  totalGradeOfHundredPassPercentWithoutDistinctionBarchartDf <- rbind(totalGradeOfHundredPassPercentWithoutDistinctionBarchartDf,
                                                                   data.frame(
                                                                     District = c(rep(d, 3)),
                                                                     Grade = c(rep("FirstDivision", 1), rep("SecondDivison",1), rep("ThirdDivision",1)),
                                                                     NoOfStudent = c(filterDistrict %>% select(FIRST.DIVISION) %>% sum(),
                                                                                     filterDistrict %>% select(SECOND.DIVISION) %>% sum(),
                                                                                     filterDistrict %>% select(THIRD.DIVISION) %>% sum())
                                                                   )
  )
}

paged_table(totalGradeOfHundredPassPercentWithoutDistinction)
```
Since, we are filtering school without having distinction. It will be obvious that there will be zero distinction. So, we can remove distinction while visualizing data.
```{r Stack bar chart representing no of student on basics of grade with zero distinction, echo=FALSE, fig.cap='Fig 6:- Total No of Student in each district on basics of grade', fig.align='c'} 
# Plot grouped barchart showing no of student in grade with 100% pass result without district kathmandu
ggplot(totalGradeOfHundredPassPercentWithoutDistinctionBarchartDf, 
       aes(fill=Grade, 
           y=NoOfStudent, 
           x=District,
           label=NoOfStudent)) + 
  geom_bar(position='dodge', stat='identity') +
  xlab('District') + 
  ylab('Total No of Student')

```
From above barchart, we can conclude that Kathmandu district had the maximum pass rate in comparision to another district. So we can assume that another district lacks the proper educational facilities as well as school.

## Max number of student within hunderd pass percent
Since, in hundred pass percent data set it is natural that the school having highest number of student is the school scoring the highest pass rate. As well as, since total no of student is not same in every school so we can't compare the grade 

```{r School with highest number of student, echo = FALSE} 
passSchoolWithMaxStudentDataFrame <- data.frame()

for (d in districts) {
  filterDistrict <- filter(hundredPercentPassPercentDataFrame, District == d)
  
  passSchoolWithMaxStudentDataFrame <- rbind(passSchoolWithMaxStudentDataFrame, filterDistrict %>% filter(PASS == max(PASS)))
  
}

paged_table(select(passSchoolWithMaxStudentDataFrame, District, Name.of.School, PASS))
```


## General summary of fail result {.tabset .tabset-fade}
As we can see, the school having fail percentage also able to secure distintion in SLC exam. So, this section will summarize the total no of student on basics of grade having zero distinction with fail percentage and vice versa. 

### Having Distinction
``` {r No of student on baiscs of grade having fail percentage school scoring distinction, echo=FALSE}
paged_table(totalGradeOfDistinctionFailPercentDataFrame)
``` 

Let's plot the above information in stack barchart for data visualiztion 
```{r Stack barchar of student on basics of grade having fail percentage school scoring distinction, echo = FALSE}
ggplot(totalGradeOfDistinctionFailPercentBarChartDf, 
       aes(fill=Grade, 
           y=NoOfStudent, 
           x=District,
           label=NoOfStudent)) + 
  geom_bar(position='dodge', stat='identity') +
  xlab('District') + 
  ylab('Total No of Student')
``` 

### Without Having Distinction 
``` {r No of student on baiscs of grade having fail percentage school without any distinction, echo=FALSE}
paged_table(totalGradeOfZeroDistinctionFailPercentDataFrame)
``` 

Since, we are filtering total no of school with zero distinction. So, it is not required to plot distinction while making barchart
```{r Stack barchar of student on basics of grade having fail percentage school without any distinction, echo = FALSE}
ggplot(totalGradeOfZeroDistinctionBarChartDf, 
       aes(fill=Grade, 
           y=NoOfStudent, 
           x=District,
           label=NoOfStudent)) + 
  geom_bar(position='dodge', stat='identity') +
  xlab('District') + 
  ylab('Total No of Student')
``` 

# Maximum number of passed student within fail percent 
Even in fail percent, we can filter the school having highest number of pass student. Like previous, since total no of student is not same in every school so we can't compare grade. The table below, summarize the information with name of school having maxium pass student.

```{r School with highest pass percent among fail percent, echo = FALSE}
failSchoolWithMaxStudentDataFrame <- data.frame()

for (d in districts) {
  filterDistrict <- filter(distinctionFailPercentDataFrame, District == d)
  
  failSchoolWithMaxStudentDataFrame <- rbind(failSchoolWithMaxStudentDataFrame, filterDistrict %>% select(District, Name.of.School, PASS, FAIL) %>% filter(PASS == max(PASS)))
}
paged_table(arrange(failSchoolWithMaxStudentDataFrame, PASS))
```


## Conclusion
Furthermore, we can also calculate among 100% pass student 
which shool have the highest distinction, first division, second division, third division, fail and maximum no of student also

we can decide which school stand on top on basics of this attribute within district

we can also filter top 5 school subject wise highest

filtering data on basics of highest pass percentage and lowest pass percentage

assuming that some of the school have equal number of student 