---
title: "Data analysis and visualization"
author: "NischalShakya"
date: "December 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clearing enviroment variables
remove(list=ls())
# Using library dplyr for data manupulation
library('dplyr')
# Using library ggplot2 
library('ggplot2')

# Set current working directory 
setwd('/home/nischal/nischalshakya15.github.io/r-assignment')

# Load data from csv file into data frame
data <- read.csv('input.csv', sep = ',', header = TRUE, stringsAsFactors = FALSE)

# Remove Zone Bagmati and Development Region beacuse it is redudant
data <- select(data, -c(Zone, Development.Region))

# Calculate fail percentage
data <- data %>% mutate('FAIL.PERCENTAGE' = 100 - data$PASS.PERCENT)

districts <- c(unique(data$District))

totalNoOfStudentDataFrame <- data.frame();

hundredPercentPassPercentDataFrame <- data.frame();
schoolGradeOfHundredPassPercentWithDistinctionDataFrame <- data.frame();
schoolGradeOfHundredPassPercentWithoutDistinctionDataFrame <- data.frame();

for (d in districts) {
  
  filterDistrict <- filter(data, District == d)
  

  totalNoOfStudentDataFrame <- rbind(totalNoOfStudentDataFrame, data.frame(District = d,
                                                                           Pass = filterDistrict %>% select(PASS) %>% sum(),
                                                                           Fail = filterDistrict %>% select(FAIL) %>% sum(),
                                                                           Total = filterDistrict %>% select(PASS,FAIL) %>% sum()))
  
  hundredPercentPassPercentDataFrame <- rbind(hundredPercentPassPercentDataFrame, 
                                              filter(data, District == d & FAIL == 0 & PASS.PERCENT == max(PASS.PERCENT)))
  
  schoolGradeOfHundredPassPercentWithoutDistinctionDataFrame <- rbind(schoolGradeOfHundredPassPercentWithoutDistinctionDataFrame, 
                                                                   filter(hundredPercentPassPercentDataFrame, District == d & DISTINCTION == 0))
  
  schoolGradeOfHundredPassPercentWithDistinctionDataFrame <- rbind(schoolGradeOfHundredPassPercentWithDistinctionDataFrame,
                                                                      filter(hundredPercentPassPercentDataFrame, District == d & DISTINCTION != 0))
}
```

## School performance for the year 2063 BS

This dataset includes the school performance data for the year 2063 BS based on the average scores and pass percentage in School Leaving Certificate (SLC) examination of Bagmati zone, an examination marking the completion of grade 10.

## General summary of column used in data set
```{r general summary of data}
str(data)
```

## Filtering unique districts of Bagmati Zone

```{r filtering unique district of Bagmati Zone}
# Get unique districts from data frame
districts <- c(unique(data$District))
```
``` {r unique district, echo = FALSE}
print(districts)
```

## Bargraph showing total no of student on basics of district
```{r Bar graph showing total no of student on basics of district}
# Plot bar graph showing total no of student in each district
print(totalNoOfStudentDataFrame)
ggplot(totalNoOfStudentDataFrame, aes(x = District, y = Total, label = Total)) + geom_bar(stat = 'identity', fill = 'steelblue', color = 'black') + geom_text(size = 3, vjust = -1) + labs(title = 'Total no of student in each district') + xlab('Districts') + ylab('Total No of Student')
```

## No of school on basics of district
``` {r No of school on basics of district}
# Count no of school on basics of district
countNoofSchoolOnBasiscOfDistrict <- count(data, data$District)
print(countNoofSchoolOnBasiscOfDistrict)

# load countNoOfSchoolOnBasicsOfDistrict in donut dataframe
donut <- data.frame(
  District=countNoofSchoolOnBasiscOfDistrict$`data$District`,
  Count=countNoofSchoolOnBasiscOfDistrict$n
)

# Compute percentages
donut$fraction <- donut$Count / sum(donut$Count)

# Compute the cumulative percentages (top of each rectangle)
donut$ymax <- cumsum(donut$fraction)

# Compute the bottom of each rectangle
donut$ymin <- c(0, head(donut$ymax, n=-1))

# Compute label position
donut$labelPosition <- (donut$ymax + donut$ymin) / 2

# Compute a good label
donut$label <- paste0(donut$Count)

# Create donut 
ggplot(donut, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=District)) +
  geom_rect() +
  geom_text( x=3.5, aes(y=labelPosition, label=label), size=3) + # x here controls label position (inner / outer)
  scale_fill_brewer(palette=3) +
  scale_color_brewer(palette=3) +
  coord_polar(theta="y") +
  xlim(c(1, 4)) +
  theme_void() +
  theme(legend.position = "right")
```

## No of school on basics of geographical region
``` {r No of school on basics of geographical region}
# Count no of school on basics of geographical region
countNoofSchoolOnBasiscOfGeographicalRegion <- count(data, data$Geographical.Region)
print(countNoofSchoolOnBasiscOfGeographicalRegion)
# Plot bar graph showing total no of student on basics of geographical region 
ggplot(countNoofSchoolOnBasiscOfGeographicalRegion, aes(x = countNoofSchoolOnBasiscOfGeographicalRegion$`data$Geographical.Region`, y = countNoofSchoolOnBasiscOfGeographicalRegion$n, label = countNoofSchoolOnBasiscOfGeographicalRegion$n)) + geom_bar(stat = 'identity', fill = 'steelblue', color = 'black', width = 0.3) + geom_text(size = 3, vjust = -1) + labs(title = 'Total no of school on basics of geographical region') + xlab('Geographical region') + ylab('Total No of School')
````

## No of school having 100 percent pass result
```{r No of school having 100 percent pass result} 
# Count no of school having 100% pass result
print(select(hundredPercentPassPercentDataFrame, District, Name.of.School))
noOfSchoolWithHundredPercentPass <- count(hundredPercentPassPercentDataFrame, hundredPercentPassPercentDataFrame$District)
# Plot bar graph showing total no of school with 100% pass result 
ggplot(noOfSchoolWithHundredPercentPass, aes(x = noOfSchoolWithHundredPercentPass$`hundredPercentPassPercentDataFrame$District`, y = noOfSchoolWithHundredPercentPass$n, label = noOfSchoolWithHundredPercentPass$n)) + geom_bar(stat = 'identity', fill = 'steelblue', color = 'black', width = 0.5) + geom_text(size = 3, vjust = -1) + labs(title = 'Total no of school with 100% pass result') + xlab('District') + ylab('Total No of School')
# It seems Rasuwa was a district which is unable to acheive 100 percent pass result
```
## No of student on baiscs of grade within 100 percent pass result school scoring non-zero distinction
```{r No of student on baiscs of grade within 100 percent pass result school scoring non-zero distinction, echo=FALSE} 
totalGradeOfHundredPassPercentWithDistinction <- data.frame()

districtsHundredPassPercentWithDistinction <- c(unique(schoolGradeOfHundredPassPercentWithDistinctionDataFrame$District))

for (d in districtsHundredPassPercentWithDistinction) {
  filterDistrict <- filter(schoolGradeOfHundredPassPercentWithDistinctionDataFrame, District == d)
  
  totalGradeOfHundredPassPercentWithDistinction <- rbind(totalGradeOfHundredPassPercentWithDistinction, data.frame(
                                                                           District = d, 
                                                                           Distinction = filterDistrict %>% select(DISTINCTION) %>% sum(),
                                                                           FirstDivision = filterDistrict %>% select(FIRST.DIVISION) %>% sum(),
                                                                           SecondDivision = filterDistrict %>% select(SECOND.DIVISION) %>% sum(),
                                                                           ThirdDivision = filterDistrict %>% select(THIRD.DIVISION) %>% sum(),
                                                                           Total = filterDistrict %>% select(DISTINCTION,FIRST.DIVISION,SECOND.DIVISION,THIRD.DIVISION) %>% sum()))
}
print(totalGradeOfHundredPassPercentWithDistinction)
``` 
From above table, we can conclude that there is not a single student scoring third division from the school with 100 percent pass rate with 0 distinction 

## No of student on baiscs of grade within 100 percent pass result school scoring with zero distinction
```{r No of student on baiscs of grade within 100 percent pass result school scoring with zero distinction, echo = FALSE} 
totalGradeOfHundredPassPercentWithoutDistinction <- data.frame()

# Filtering unique district of hundred pass percent without distinction 
districtsHundredPassPercentWithoutDistinction <- c(unique(schoolGradeOfHundredPassPercentWithoutDistinctionDataFrame$District))

for (d in districtsHundredPassPercentWithoutDistinction) {
  filterDistrict <- filter(schoolGradeOfHundredPassPercentWithoutDistinctionDataFrame, District == d)
  
  totalGradeOfHundredPassPercentWithoutDistinction <- rbind(totalGradeOfHundredPassPercentWithoutDistinction, data.frame(
                                                                           District = d, 
                                                                           Distinction = filterDistrict %>% select(DISTINCTION) %>% sum(),
                                                                           FirstDivision = filterDistrict %>% select(FIRST.DIVISION) %>% sum(),
                                                                           SecondDivision = filterDistrict %>% select(SECOND.DIVISION) %>% sum(),
                                                                           ThirdDivision = filterDistrict %>% select(THIRD.DIVISION) %>% sum(),
                                                                           Total = filterDistrict %>% select(DISTINCTION,FIRST.DIVISION,SECOND.DIVISION,THIRD.DIVISION) %>% sum()))
}
print(totalGradeOfHundredPassPercentWithoutDistinction)
```

## General summary of data on basic of district
Despite of 100% pass result, it seems in some school there is zero distiction. 
So, this seciton will summarize the school on basics of district having zero distinction within 100% pass result and vice versa.

Furthermore, we can also calculate among 100% pass student which shool have the highest distinction, first division, second division, third division, fail also

we can decide which school stand on top on basics of this attribute within district

we can also filter top 5 school subject wise highest

filtering data on basics of highest pass percentage and lowest pass percentage

thinking that some of the school have equal number of student 

it seems school having distinction within 100% pass result have the 0 no of student scoring third division 
and with 0 distinction the stdudent scoring 1st division is more with less third division